---
title: "WP9 Solutions"
output: pdf_document
---

---
title: "WP9"
date: 'Due Date: 11/12/21'
urlcolor: blue
output: pdf_document
---

```{r setup, include=FALSE}
library(formatR)
library(tidyverse)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

```

1. Load the `rdrobust` package. 

```{r, message = F, warning = F}
library(rdrobust)
library(readr)

```

2. Load in the `senate.csv` dataset. Make a new dataset called `df` that has three variables: 

  - Y: made from demvoteshfor2 
  - Z: made from demmv 
  - D: a logical variable stating whether Z is greater than or equal to 0 

```{r, message = F, warning = F}
df <- read_csv("senate.csv") %>% 
  select(Y = demvoteshfor2, 
         Z = demmv)%>%
  mutate(D = Z >= 0)


```


3. Construct an RD plot of the Design. On the Y axis, plot the outcome variable. On the X axis, plot the score variable. Give this plot the title "RD Plot for Senate Elections Data" 

```{r}
## Raw comparison of means 
rdplot(df$Y, df$Z)

```

```{r}
## local comparison of means 
rdplot(df$Y[abs(df$Z) <= 50], df$Z[abs(df$Z) <= 50])

```


4. Construct an alternative RD plot with evenly spaced bins. Give this plot the same title and use the same variables. 

```{r}
# One way to do it without messing around with the number of bins 
rdplot(df$Y, df$Z, binselect = "es")

```

5. Conduct an RD robust estimate with default options. Report the output and analyze whether there is a statistically significant effect. 

```{r}
default <- rdrobust(df$Y, df$Z)
summary(default)
```


6. Conduct an RD robust estimate with uniform weights and with triangular weights. Report both estimates. You will need to look at the options of the canonical function of `rdrobust` to do this. 

```{r}
# uniform weights 
uniform <- rdrobust(df$Y, df$Z, kernel = "uniform")
summary(uniform)
```

```{r}
# triangular weights 
triangle <- rdrobust(df$Y, df$Z, kernel = "triangular")
summary(triangle)

```

7. Do you think there is an effect of treatment in this RDD? If so what it is? What are the external validity limitations inherent in this design? 

*Yes, there is a treatment effect by our estimates. It's a strong treatment effect as well. The inherent limitations is that for any RDD is that we can only get a local effect. It is not clear that a design tells us anything about the general estimand of interest because RDD identify a LATE.*

