---
title: "Weekly Practice 3"
author: "FILL IN YOUR NAME"
output: pdf_document 
---

## Pedagogical Purpose 

The point of this Weekly practice is to walk through the CACE Theorem. 

### Part A 

Build a data frame according to the following specifications 

N <number> a column of units. There should be 100. 

YiD0 <number> a potential outcome under control. This should be the same length as N from a standard normal distribution. 

YiD1 <number> a potential outcome under treatment. This should be a constant effect of 5. 

diZ0 <number> This is the potential outcome given treatment is never received. For all units, it should be the same number that we use to represent no treatment. 

diZ1 <number> This is the status for units who are compliers. Use complete random assignment and put 30% of units in treatment. 

Type <number> This is a dummy variable indicating whether a unit is a complier or a NT. If a unit is a complier, this is equal to 1, else this is equal to 0. 

```{r, message = F, warning=F}
library(tidyverse)
set.seed(42)
data <- tibble(
  N = 1:100,
  YiD0 = rnorm(100),
  YiD1 = YiD0 + 5,
  diZ0 = c(rep(0,100)),
  diZ1 = sample(c(rep(0,30), rep(1,70)), 100, F),
  Type = if_else(diZ1 == 1, 1, 0)
)
```


### Part B 

Identify the $ITT_D$ which is $E[d_i(z=1) - d_i(z=0)]$. 

```{r}


```


Identify the $ITT$ which is $E[Y_i(z=1, d(1))] - E[Y_i(z=0, d(0))]$


```{r}


```

### Part C 

Identify the CACE which is the ratio of the $ITT$ relative to the $ITT_D$

```{r}


```

### Part D 

Turn your work from parts B and C into a function called CACE. This function should take in a data frame with appropriate columns and return the CACE. 

Show your function works by passing the created dataset. You should get the same answer as Part C if your function is correct. 