---
title: "WP6 Solution" 
output: pdf_document
---


```{r setup, include=FALSE}
library(formatR)
library(tidyverse)
library(estimatr)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

```


## Read in data 

```{r, message=F, warning = F}
segsub <- read_csv("Segregation_HS.csv")%>%
  filter(seg == 1)

int <- read_csv("Integration_HS.csv")
```

## Figure 2 

We need to do a bit of data wrangling to get this answer 

First, blackbn is a number. We should turn it into a character. 

```{r}
segsub <- segsub %>% 
  mutate(blackbn = as.character(blackbn))
```

Next, we need to group by the count and blackbn variables. Within each of these groupings we want to get the total number (`n()`) and the average fatality rate. The variable for this from the replication codebook is `spec_fatal_rate`

```{r}
fig2Data <- segsub%>% 
  group_by(count, blackbn)%>%
  summarise(number = n(), 
            sum = sum(spec_fatal, na.rm = T),
            avg = mean(spec_fatal_rate, na.rm = T)*100, .groups = "drop_last")
```

Now that we have the data, we can make the figure. 

```{r}
segsub %>% 
  ggplot(aes(x = count, y = spec_fatal_rate100, shape = blackbn))+
  geom_point()+
  geom_line(data = fig2Data, aes(x = count, y = avg, group = blackbn, color = blackbn))+
  xlab("Period (Year-Month-Half)")+
  ylab("Fatality (%)")+
  theme(legend.title = element_blank())

```

## Figure 3 


We follow the same steps as before but with the int data set. 

```{r}
int <- int %>% 
  mutate(blackbn = as.character(blackbn))


fig3Data <- int %>% 
  group_by(count, blackbn)%>%
  summarise(number = n(), 
            avg = mean(specpct_imputed, na.rm = T)*100,
            sum = sum(specpct_imputed, na.rm = T),
            .groups = "drop_last")

## Figure 3 
ggplot(data = int, aes(x = count, y = specpct_imputed*100, shape = blackbn))+
  geom_point()+
  geom_line(data = fig3Data, aes(x = count, y = avg, group = blackbn, color = blackbn), size = 1)+
  xlab("Period (Year-Month-Half)")+
  ylab("Fatality (%)")+
  theme(legend.title = element_blank())
```


## Replicate OLS models for Table 1

To replicate the models, we look to see the appropriate variables, and then make the coefficient plot as usual. 

The one "new" thing here is fixed effects, which we will cover next week. 

```{r}
m1 <- lm_robust(spec_fatal_rate100 ~ blackbn, data = segsub, se_type = "HC3")%>%
  tidy()%>%
  mutate(model = "m1")

m2 <- lm_robust(spec_fatal_rate100 ~ blackbn, data = segsub, fixed_effects = ~count,
                se_type = "HC3")%>%
  tidy()%>%
  mutate(model = "m2")

## Coefficient plot of models for Table 1 
cplot <- bind_rows(m1, m2)%>%
  filter(term == "blackbn1")%>%
  ggplot(aes(x = estimate, y = model, color = model))+
  geom_point()+
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, height = 0.1))+
  theme_bw()+
  guides(color = "none")+
  ggtitle("Coefficient Plots for Table 1")
cplot
```

## Replicate Table 3

Conceptually, nothing is different than what was required for the replication of Table 1. 

```{r}
m31 <- lm_robust(specpct_imputed100 ~ blackbn, data = int, se_type = "HC3")%>%
  tidy()%>%
  mutate(model = "m1")
m32 <- lm_robust(specpct_imputed100 ~ blackbn, data = int, fixed_effects = ~count,  
                 se_type = "HC3")%>%
  tidy()%>%
  mutate(model = "m2")
m33 <- lm_robust(specpct_imputed100 ~ blackbn, data = int, fixed_effects = ~count + obs, se_type = "HC3")%>%
  tidy()%>%
  mutate(model = "m3")

cplot2 <- bind_rows(m31, m32, m33)%>%
  filter(term == "blackbn1")%>%
  ggplot(aes(x = estimate, y = model, color = model))+
  geom_point()+
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, height = 0.1))+
  theme_bw()+
  guides(color = "none")+
  ggtitle("Coefficient Plots for Table 3")
cplot2
```
