---
title: "Weekly Practice 1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
```

## Problems 

### Libraries 

1. Load the `tidyverse` package. 

```{r, eval = F, echo = F}
library(tidyverse)
```

2. Provide the code needed to install the `estimatr` package and the `here` package. You should also install this package on your machine for the future. 

```{r, echo = F, eval = F}
install.packages(c("estimatr", "here"))

```

### Plotting Practice 

On bCourses is a csv file called `wpData.csv` 

**Question 1**

a. Read in this file to R as a tibble. Name your tibble "img" (without quotes).  

```{r, echo = F, eval = F}
img <- read_csv(here("data/wpData.csv"))

```

b. Using `ggplot2` make a scatterplot of the data. Set the color aesthetic to the z column. In your chain add the functions `scale_color_identity()` and `theme_void()`. Using the function to give your plot a title, set the title to "How Learning Often Feels." (Note, this may take a little bit of time to run).

```{r, eval = F, echo = F}
ggplot(img)+
  geom_point(mapping = aes(x=x, y = y, color = z))+
  scale_color_identity()+
  theme_void()+
  ggtitle("How Learning Often Feels")

```

**Question 2**

a. Set the RNG seed to 42, and run it to make sure that your random number generator starts in the right place. 

b. Create a tibble named `practice` with five variables: 

- id: <numeric vector> the sequential set of numbers beginning at 100 and ending at 200 inclusive. 

- x: <numeric vector> a set of observations from a uniform distribution with a minimum of 7 and a maximum of 42 

- y: <numeric vector> a set of observations from a standard normal distribution 

- u: <numeric vector> a set of observation from a normal distribution centered at 100 with variance 100 

- d: <numeric vector> a set of binomial observations from a fair coin. 

c. What is the number of rows of the data frame you made? What is the number of columns?

d. Use a function to print the first six rows of your data frame. 

```{r, echo = F, eval = F}
set.seed(42)

practice <- tibble(
  id = 100:200,
  x = runif(101, 7, 42),
  y = rnorm(101, 0,1),
  u = rnorm(101, 100, 10),
  d = rbinom(101, 1, c(0.5,0.5))
)

head(practice)
```

\pagebreak 

**Question 3** 

For each of the following code blocks, explain why will result in an error. Fix the error to make the code run. 

a. 

```{r, echo = T, eval = F}
A_VARIABLE <- 25
A_VAR1ABLE 
```

b. 

```{r, eval = F}
multNum <- function(arg1, arg2){
  out <- arg1 * arg2
  return(out)
}

multNum(2) == 10

```

c.

```{r, eval = F}
myTibble <- tibble(
  x = c(1,2,4,6),
  y = c("John", "Paul", "George, "Ringo"),
  z = c(TRUE, FALSE, TRUE, FALSE)
) 
```


