---
title: "Weekly Practice 1"
author: "FILL IN YOUR NAME"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
```

## Delete me after reading 

Below is a template for Weekly Practice 1. Make sure to run all of your code when you feel like it is correct. When finished, knit your file to a pdf and submit the pdf. 

## Problems 

### Libraries 

1. Load the `tidyverse` package. 

```{r}
### Fill in your answer here 
### Run the chunk when finished 
```

2. Provide the code needed to install the `estimatr` package and the `here` package. You should also install this package on your machine for the future. 

```{r}
### Fill in your answer here 

```

### Plotting Practice 

On bCourses is a csv file called `wpData.csv` 

**Question 1**

a. Read in this file to R as a tibble. Name your tibble "img" (without quotes).  

```{r}
img <- ... # fill in your answer here, replacing the ... 

```

b. Using `ggplot2` make a scatterplot of the data. Set the color aesthetic to the z column. In your chain add the functions `scale_color_identity()` and `theme_void()`. Using the function to give your plot a title, set the title to "How Learning Often Feels." (Note, this may take a little bit of time to run).

```{r}
### Fill in your answer here 

```

**Question 2**

a. Set the RNG seed to 42, and run it to make sure that your random number generator starts in the right place. 

b. Create a tibble named `practice` with five variables: 

- id: <numeric vector> the sequential set of numbers beginning at 100 and ending at 200 inclusive. 

- x: <numeric vector> a set of observations from a uniform distribution with a minimum of 7 and a maximum of 42 

- y: <numeric vector> a set of observations from a standard normal distribution 

- u: <numeric vector> a set of observation from a normal distribution centered at 100 with variance 100 

- d: <numeric vector> a set of binomial observations from a fair coin. 

c. What is the number of rows of the data frame you made? What is the number of columns?

d. Use a function to print the first six rows of your data frame. 

```{r}
set.seed(...) # Replace the ... with your answer 

### Fill in the rest of your answer here 


```


**Question 3** 

For each of the following code blocks, explain why will result in an error. Fix the error to make the code run. 

a. 

```{r}
## Fix this code 
A_VARIABLE <- 25
A_VAR1ABLE 
```


b. 

```{r}
multNum <- function(arg1, arg2){
  out <- arg1 * arg2
  return(out)
}

# Fix this code 
multNum(2) == 10

```

c.

```{r}
# Fix this code
myTibble <- tibble(
  x = c(1,2,4,6),
  y = c("John", "Paul", "George, "Ringo"),
  z = c(TRUE, FALSE, TRUE, FALSE)
) 
```


