---
title: "Applied Causal Inference"
subtitle: "Lecture 1"  
author: 
  - Alex Stephenson
date: '8-25-2021'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#003262",
                 base_font_size = "30px")

```

## What are we doing today?

1. Course Overview 
2. Introductions 
3. Take the Final Exam

---
class: highlight-last-item 
## Course Overview 

We have three goals in this course 

- Understand the concepts of applied causal inference 



- Demonstrate our learning through code that we can take to future research or jobs 



- Enjoy doing Goals 1 and 2



My job is to help you learn this material as best as possible. Think of me as a coach. 

---

## Where do I find stuff? 

- The syllabus has all basic information for the course. 



- I cannot emphasize enough that you should read the syllabus 



- Seriously, instructors spend lots of time on the syllabus. Humor us by reading the syllabus. 
---

## Yeah, but what if it's not in the syllabus?

- Feel free to reach out to me or your fellow classmates. 


- Lots of information comes through email and bCourses. Check them frequently. 
---

class: center, middle 
## Course Mechanics 

---
class: highlight-last-item 
## Lecture Mechanics 

Berkeley time means we 150 minutes of lecture a week. I plan to split up the lectures as follows: 

- First 50 Minutes: Concept overview 



- Second 50 Minutes: Basic application of the concept 



- Third 50 Minutes: Additional application of the concept. 

- Lecture slides go up on bCourses after lecture.
---
## What should I bring to lecture?

- On Mondays: 
  - Whatever you like to take notes on. 



- On Wednesdays and Fridays: 

  - Whatever notes you thought were useful
  - Anything else that you think will be helpful, but at minimum a computer. 

---

## A Brief Digression on Notes and Studying 

- You're probably taking notes wrong 

- You're probably studying wrong 

- If either or both are true, it's not your fault. No one taught you. 

---
## Section Mechanics 

Section is R heavy by design. 

The purpose of section is to prepare you for applications and assignments. 

You will always need a laptop in Section. 

Section materials will always be posted on bCourses, often before section begins.

---

## Assignments 

- This class works differently than other classes you have taken at Berkeley. 

- We're working together during a pandemic, so flexibility to changing circumstances is a virtue. 

---

class: center, middle 
## Introductions 

---
class: center, middle
## Final Exam
---

## Are you \*\*\*\*\*\*\* kidding?

---
## No, but it's short. 

The answers themselves have no effect on your grade. It is for you and I to have a record of where you started at the beginning of class. 

Completing and turning in the assignment counts as the first Checkpoint. 

We will cover all the material on the final exam over the semester. 

Please turn it in before you leave today. I'll hand it back in section. 