---
title: "Regression"
author: "Alex Stephenson"
date: "5/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, message = FALSE, warning=FALSE}
library(tidyverse)
library(patchwork)
```

```{r}
X <- c(1.37, 0.3, 1.57, -.69, -.64, .08, .47, 1.33, .13, .59, -.44, .76, .73, -3.54, -0.05)

Y <- c(6.88, 1.17, 16.16, -4.84, -6.17, 0.67, 3.36, -.69, 8.03, -3.35, 1.87, 13.54, 4.06, 6.72, 7.20)

X_real <- seq(-4,4,by=c(8/14))
CEF <- 1 + 2*X_real
df <- tibble(X = X,
             Y = Y,
             CEF = CEF,
             X_real = X_real)

```

Graph CEF, points, and regression line 

```{r, message = FALSE}
df %>%
  ggplot(aes(X_real, Y))+
  geom_point(aes(x=X, y = Y))+
  geom_line(aes(y=CEF))+
  geom_smooth(method = "lm", se = F, color = "red")+
  ylim(-20,20)
```

Note that while not exactly correct, the two are reasonably close. Suppose we had 1000 observations. 

```{r}
X <- rnorm(1000)
U <- rnorm(1000)
Y <- 1 + 2*X + 5*U
CEF <- 1 + 2*X
df <- tibble(X = X,
             Y = Y,
             CEF = CEF)

df %>%
  ggplot(aes(X, Y))+
  geom_point(aes(x=X, y = Y))+
  geom_line(aes(y=CEF), color = "blue")+
  geom_smooth(method = "lm", se = F, color = "red")+
  ylim(-20,20)
```

The lines are basically on top of each other. Since the CEF is linear, the BLP is the CEF. 

## Best Quadratic Predictor 

Suppose the same setup as the micronumerosity example. 

```{r}
X <- runif(1000)
Y <- 10*X^2 + rnorm(1000)
CEF <- 10*X^2 
BLP <- -5/3 + 10*X

quad <- tibble(
  X = X,
  Y = Y,
  CEF = CEF,
  BLP = BLP
)

quad %>% 
  ggplot(aes(X,BLP))+
  geom_line(color = "red")+
  geom_line(aes(x = X,y=CEF), color = "black")+
  ylab("Y")
```
Not bad, but we might be able to do better if we could make a curved BLP. Fortunately, we can add a new explanatory predictor $X^2$ and plug that into the model. 

$$\begin{aligned}
Y &= \beta_0 + \beta_1X + \beta_2X^2
\end{aligned}$$

In this example, the BLP of $Y|X,X^2$ is 

$$\begin{aligned}
g(X,X^2) &= E[Y|X,X^2]\\
g(X,X^2) &= E[Y|X]\\
g(X,X^2) &= \beta_0+ \beta_1X + \beta_2X^2 \\
g(X,X^2) &= 0 + 0X + 10X^2 \\
g(X,X^2) &= 10X^2
\end{aligned}$$

Once again if the CEF is linear in this specification, then the BLP is the CEF. To interpret the coefficients, we find the slope of the line of a nonlinear function by taking the partial derivative of the CEF with respect to X. 

$$\begin{aligned}
\frac{\partial E[Y|X]}{\partial X} &= \frac{\beta_0 + \beta_1X + \beta_2X^2}{\partial X}\\
\frac{\partial E[Y|X]}{\partial X} &= \beta_1 + 2\beta_2X
\end{aligned}$$

In general as long as the CEF is continuous and the support of X is on the interval, we can use polynomials to get the MSE as small as we'd like. Here is another example. 

```{r}
# TODO Fill in this example 
cef <- function(x){
  # A function to create a nonlinear CEF 
  # x = a numeric value
  if(x >= -1 & x <=0){
    return(0)
  }

  if(x > 0 & x < 1){
    return(x^2)
  }
  if(x >=1 & x <= 3){
    return(2-x)
  }
}

x <- seq(-1,3, by = 0.01)
CEF <- unlist(lapply(x, cef))

df <- tibble(
  x = x,
  CEF = CEF, 
)

df %>% 
  ggplot(aes(x, CEF))+
  geom_line()
```

## Overfitting Example 

```{r, message = F, warning = F}
overfit_plot <- function(N){
  set.seed(42)
  p<-tibble(
    x = runif(N, -10, 10),
    e = rnorm(N, 0, 25),
    y = x + 5 + e,
    CEF = x + 5)%>% 
    ggplot(aes(x = x, y))+
    geom_point(aes(y=y))+
    geom_line(aes(y=CEF,color = "CEF"), size = 1)+
    stat_smooth(aes(color = "Linear Fit"),method = "lm", formula = y~x, size = 1, se = F)+
    stat_smooth(aes(color = "5th Order Polynomial"),method = "lm", formula = y~poly(x, 5, raw = TRUE), 
                size = 1,se = F)+
    ggthemes::theme_clean()+
    scale_color_manual(name = "Fit", 
                     values = c("CEF" = "black", "Linear Fit" = "blue", "5th Order Polynomial"="red"),
                     guide = "legend")+
    theme(legend.position = "bottom")+
    ylim(-45,100)+
    ggtitle(paste0("Fit with N = ", N))
  p
}

plots <- map(list(10,100,1000), ~overfit_plot(.x))

plots[[1]] + (plots[[2]] / plots[[3]])+
  plot_layout(widths = c(4,4))

```

