---
title: "Weekly Practice 2"
author: "FILL IN YOUR NAME HERE" 
output: pdf_document 
---

## Omitted Variable Bias 

The WP for this week is an example of what happens when we have omitted variable bias. 

## Part A 

Load the appropriate libraries needed to be able to appropriately estimate a linear model in R, and then graph the coefficients. 


```{r, warning = F, message = F}
# The chunk here has these settings on purpose. They will prevent the packages you load from generating messages that are irrelevant to the exercise
```


## Part B 

Build a simulated dataset. Set the random seed for reproducibility to 42. 

Imagine your dataset has 1000 observations. It should contain the following: 

D: a treatment vector that can take on either the values of 0 or 1. 
X: a covariate that affects Y that is normally distributed with a mean of 100 and a standard deviation of 50
Y_short: a variable that is the result of adding the treatment assignment and a random standard normal draw. 
Y_long: a variable that is the result of adding the treatment assignment, X, and a random standard normal draw. 

```{r}


```


## Part C 

Run the result of the regression of Y_short on D. Get the output as a data frame and save the result to the variable m1. Add a column named model to m1 that is defined for every variable as your formula call. 

```{r}


```


Run the result of the regression of Y_short on D and X. Get the output as a data frame and save the result to the variable m2.Add a column named model to m2 that is defined for every variable as your formula call. 

```{r}


```

Join the two data frames together. Call this data frame result1. Consult the `bind_rows()` function from dplyr for a way to do this. Graph just the D terms for each model on the same graph. You do not need to include error bars. 

```{r}


```

## Part D 

Now repeat the same exercise as Part C, but using Y_long as the outcome variable. 

Run the result of the regression of Y_long on D. Get the output as a data frame and save the result to the variable m3. Add a column named model to m3 that is defined for every variable as your formula call. 

```{r}


```


Run the result of the regression of Y_long on D and X. Get the output as a data frame and save the result to the variable m4. Add a column named model to m4 that is defined for every variable as your formula call. 

```{r}


```

Join the two data frames together. Call this data frame result2. Consult the `bind_rows()` function from dplyr for a way to do this. Graph just the D terms for each model on the same graph. It is not required to put confidence intervals around your points. 

```{r}


```

## Part E 

Based on the two graphs you made, what is a conclusion about the effect of omitted variables? 

