---
title: "Section 7"
date: "10-20-2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(tidyverse)
library(knitr)
library(sensemakr)
library(estimatr)
style_mono_light(base_color = "#003262",
                 base_font_size = "25px")
```

## Sensitivity Analysis in Practice 

We spent time on sensitivity analysis in lecture. 

Today we're going to see an approach with software in R called `sensemakr` 

---

## What Question are we answering?

How strong would an unobserved confounder (or a group of confounders) have to be change a research conclusion?

In a worst-case scenario, how robust our our results to all unobserved confounders acting together?

How strong would confounding need to be relative to the strength of observed covariates to change the answer by a certain amount?

---

## Why does this matter?

First we should more or less never expect there is no unobserved confounders. 

As a consequence, we always have to make an argument that what is left out is trivial in an observational study. 

Sensitivity analysis provides a quantitative way to discuss the fragility of an estimate.

???

Incidentally, we have to do this too in a regular experiment, but our design gives us substantial plausibility the gravitational effect of Neptune probably isn't doing much. 

There's a practical implication to this. Chaudoin, Hays, and Hicks (2018) replicate a set of studies looking at the effect of membership in the WTO. They find that over a third of those regressions produce an effect of membership on cancer rates. 

---

## sensemakr    

`sensemakr` is an R package that implements a particular type of sensitivity analysis based on an omitted variable bias framework. 

Note that there are other sensitivity frameworks. Here we are considering the effect of unobserved variables. 

Other methods like we discussed in lecture we can usually implement from scratch. 

---

## Motivating Example: Violence in Darfur 

The data comes from a survey 

"Did direct exposure to violence make individuals more likely to ask for peace or ask for revenge?" 

???

In 2003 and 2004, the Darfurian government orchestrated a horrific campaign of violence against civilians, killing an estimated two hundred thousand people. This application asks whether, on average, being directly injured or maimed in this episode made individuals more likely to feel “vengeful” and unwilling to make peace with those who perpetrated this violence. Or, might those who directly suffered such violence be motivated to see it end, supporting calls for peace?

Assumption: government bombings and attacks by militias known as the Janjaweed were indiscriminate within village, except that gender will have a possible differential effect due to sexual violence targeted towards women 
---

## The Argued Design 

```{r, eval = F}
lm_robust(peacefactor ~ directlyharmed + female, 
          fixed_effects = village, 
          data = darfur)

```

```{r, echo = F}
data("darfur")
dm <- lm_robust(peacefactor ~ directlyharmed + female, fixed_effects = village, data = darfur)%>% 
  tidy()
knitr::kable(dm)
```

???

What's the problem here?

Well there are lots of possible other variables we could have looked at. What about differential effects of bombings? Bombs might be more likely to be targeted at the centers of villages. People do not live in villages at random and so individuals closer to the center may be affected differently than individuals on the periphery. 

Further, perhaps support for peace is also based on social position. Wealth and political attitudes might affect whether someone supports peace. We could imagine that these are moderators, such that different values imply different sub groups. There might be interaction effects. Maybe rich people support peace even more because they're rich. 

Of course, like any omitted variable these are not measured, what would have been the effect if they were?
---

## Assessing the Design 

```{r, eval = F}
dm <- lm(peacefactor ~ directlyharmed + female + village, data = darfur)
dm.sense <- sensemakr(model = dm, 
                      treatment = "directlyharmed",
                      benchmark_covariates = "female",
                      kd = 1:3)
ovb_minimal_reporting(dm.sense)
```

???

Here we need a model. For this package, it's a lm() object 

Second we need to tell what the treatment is 

Third we need the covariates to bound plausible confounders. 

Fourth we tell sensemakr how many more times stronger the confounder is to the treatment in explaining outcomes 
---
## Assessing the Design

```{r, echo =F, results='asis'}
dm <- lm(peacefactor ~ directlyharmed + female + village, data = darfur)
dm.sense <- sensemakr(model = dm, 
                      treatment = "directlyharmed",
                      benchmark_covariates = "female",
                      kd = 1:3)
ovb_minimal_reporting(dm.sense, format = "html")
```

???

Partial R^2 of the treatment with the outcomes tells us that in an extreme scenario even if confounders explained all remaining variation of the treatment, they would need to explain at least 2.2% of the residual variation of the treatment to bring down the estimated effect to zero. 

Note that this might be very little or a lot depending on your application. Putting a number on something does not absolve us from having to think 

Robustness value: (RV) if confounders explain 13.9% of both the residual variation of the outcome and of the treatment this is sufficient to explain away the effect. At least one of them greater than 13.9% is necessary, otherwise such confounders cannot explain away the point estimate. 


---
