---
title: "Attrition Example"

---

```{r}
options(htmltools.dir.version = FALSE)
library(tidyverse)

```

## Data

```{r}
data <- tibble(
  unit = 1:8,
  y0 = c(3,3,8,7,6,5,6,6),
  y1 = c(8,7,10, 8,6,8,6,7),
  r0 = c(0,0,1,1,0,1,1,1),
  r1 = c(1,1,1,1,0,1,0,1),
  y0_r0 = c(NA, NA, 8,7, NA, 5,6,6),
  y1_r1 = c(8,7,10,8,NA, 8, NA, 7)
)

```

## What is the attrition here 

In this data, subject 5 is always missing. Subjects 1,2 are missing if they are in the control arm, and Subject 7 is missing if they are in the treatment arm. 

## ATE 

Oski the Bear land lets us just compare potential outcomes to get the true value here. 

```{r}
data %>% 
  summarise(ATE = mean(y1) - mean(y0))%>%
  pull()

```

The True ATE is 2. An unbiased estimator should give us this number back. 


## Comparing Non-Missing Values 

```{r}
data %>% 
  summarise(ATE = mean(y1_r1, na.rm = T)- mean(y0_r0, na.rm = T))%>%
  pull()

```

If we randomly assign units to treatment and control and compare average observed Y values, we would find on average a difference of 1.6, not 2. Hence this procedure is biased when attrition is not random. 

Here attrition is correlated with the control group. More subjects want to take the treatment and more precisely lower values of $Y_i(z)$ are missing. 


## ATE Generalizing 

The ATE can be defined as 

$$E[R_i(1)]E[Y_i(1)|R_i(1)=1] + (1-E[R_i(1)])E[Y_i(1)|R_i(1)=0]) - \\
E[R_i(0)]E[Y_i(0)|R_i(0)=1] - (1 - E[R_i(0)])E[Y_i(0)|R_i(0)=0])$$ 

Let's calculate this equation with our data 

```{r}
## First part of equation for treatment group 
r1 = mean(data$r1)
y1r1 = mean(data$y1_r1, na.rm = T) # Only observed 
r1_complement = 1 - r1
y1r10 = data %>% 
  filter(is.na(y1_r1))%>%
  summarise(avg = mean(y1))%>%
  pull()
```


```{r}
## Second part for control group 
r0 = mean(data$r0)
y0r0 = mean(data$y0_r0, na.rm = T) # Only observed 
r0_complement= 1 - r0
y0r00 = data %>% 
  filter(is.na(y0_r0))%>%
  summarise(avg = mean(y0))%>%
  pull()
```

We see that if we know everything about the problem, we can get the appropriate weights back. However, if we only view the observed values then we will get a different value. 

```{r}
# put it all together 
part1 = r1 * y1r1 + r1_complement*y1r10 
part2 = r0 *y0r0 + r0_complement*y0r00
part1 - part2
```

